---
title: "Practice Dec 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache= TRUE, autodep = TRUE)

library(rethinking)
library(ggplot2)
library(reshape2)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

## Problem 1 : Tomato

Remember the tomato data set generated by Pepe; we first looked at this when we were working on Chapter 9.  35 accessions for seven species were grown in sun and shade.  

Assess whether there is evidence for total length ("totleng") response to shade and whether this response varies by species.  Consider whether including accession ("acs"), and/or shelf ("shelf") using adaptive priors improves the model fit.

Bonus: would it be better to consider shade by accession interactions instead of shade x species?
```{r}
getwd()
data <- read.csv("/Users/xyyue/Documents/RClub/Rclub-rethinking_Xiaoyan.Yue/Assignment_Chapter_09/TomatoR2CSHL.csv")

#take a look at the data
head(data)
summary(data)
str(data)

#categorical variable for trt
levels(data$trt)
data$trt2 <- as.numeric(data$trt)-1 #0 = H, 1=L

#categorical variables for species
#almost the same thing as the last step, set the dummy variable "species" into different column, which is really useful when you have a lot of species, you can not do it one by one
data$index <- 1:nrow(data)
data2 <- dcast(data,index + totleng + trt2 ~ species, value.var = "species",fun.aggregate = length)

#fix species names to get rid of space 
colnames(data2) <- sub(". ","_",colnames(data2))
head(data2)
summary(data2)

#set a model with a prior distribution using map2stan
##stan needs a "clear data", we must subset the data frame to contain only the relevant columns and get rid of the other variables
data2.trt <- data2[,c("totleng","trt2")]

#First, fit a model with totleng and trt
#set the model for totleng and trt
trt.stan <- map2stan(alist(
  totleng ~ dnorm(mu,sigma),
  mu <- a + bT * trt2,
  a ~ dnorm(53.7,100),
  bT ~ dnorm(0,10),
  sigma ~ dunif(0,20)),
  data2.trt,
  chains = 4)

#test the model
precis(trt.stan)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(trt.stan))

#let's get the posterior distribution of bT, the treatment coefficient.
post.bT <- extract.samples(trt.stan)$bT
dens(post.bT,show.HPDI = 0.95)
#so, the trt do has a positive effect on totleng

# is there any percent of the posterior distribution of bT is less than or equal to 0?
sum(post.bT <= 0) / length(post.bT) 
# it turns out None of the posterior distribution for bT is less than or equal to 0.

#Next,fit a model with totlength,trt and species
data2.species.trt <- data2[,c(2:8)]
head(data2.species.trt)

#set the model for totleng, trt and species
species.trt.stan <- map2stan(alist(
  totleng_log ~ dnorm(mu,sigma),
  mu <- a + bT*trt2 + bChil*S_chilense + bChmi*S_chmielewskii + bHab*S_habrochaites + bPen * S_pennellii + bPer*S_peruvianum,
  a ~ dnorm(5.7,10),
  c(bT,bChil,bChmi,bHab,bPen,bPer) ~ dnorm(0,10),
  sigma ~ dunif(0,20)),
  data2.species.trt,
  chains = 4)

plot(species.trt.stan)
precis(species.trt.stan)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(species.trt.stan))

#compare the models
compare(trt.stan,species.trt.stan)

#consider accession
str(data)
data3 <- dcast(data,index + totleng_log + trt2 + acs +shelf ~ species, value.var = "species",fun.aggregate = length)

head(data3)

#fix species names to get rid of space 
colnames(data3) <- sub(". ","_",colnames(data3))

data3$acs <- coerce_index(data3$acs)
data3$shelf <- coerce_index(data3$shelf)

#fit a model including accession ("acs"), and/or shelf ("shelf") using adaptive priors 
data3.acs <- data3[,c(2:4,6:10)]
acs.stan <- map2stan(alist(
  totleng_log ~ dnorm(mu,sigma),
  mu <- a + bT*trt2 + bChil*S_chilense + bChmi*S_chmielewskii + bHab*S_habrochaites + bPen * S_pennellii + bPer*S_peruvianum + a_acs[acs],
  a ~ dnorm(5.7,10),
  c(bT,bChil,bChmi,bHab,bPen,bPer) ~ dnorm(0,10),
  sigma ~ dunif(0,20),
  a_acs[acs] ~ dnorm(0,sigma_acs),
  sigma_acs ~ dcauchy(0,1)),
  data3.acs,
  chains = 4)

precis(acs.stan,depth=2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(acs.stan))

#fit a model including shelf using adaptive prior
data3.shelf <- data3[,c(2:3,5:10)]
shelf.stan <- map2stan(alist(
  totleng_log ~ dnorm(mu,sigma),
  mu <- a + bT*trt2 + bChil*S_chilense + bChmi*S_chmielewskii + bHab*S_habrochaites + bPen * S_pennellii + bPer*S_peruvianum + a_shelf[shelf] ,
  a ~ dnorm(5.7,10),
  c(bT,bChil,bChmi,bHab,bPen,bPer) ~ dnorm(0,10),
  sigma ~ dunif(0,20),
  a_shelf[shelf] ~ dnorm(0,sigma_shelf),
  sigma_shelf ~ dcauchy(0,1)),
  data3.shelf,
  chains = 4)

precis(shelf.stan,depth=2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(shelf.stan))

#fit a model including acs and shelf using adaptive priors
data3.shelf <- data3[,c(2:3,5:10)]
shelf.acs.stan <- map2stan(alist(
  totleng_log ~ dnorm(mu,sigma),
  mu <- a + bT*trt2 + bChil*S_chilense + bChmi*S_chmielewskii + bHab*S_habrochaites + bPen * S_pennellii + bPer*S_peruvianum + a_shelf[shelf] + a_acs[acs],
  a ~ dnorm(5.7,10),
  c(bT,bChil,bChmi,bHab,bPen,bPer) ~ dnorm(0,10),
  sigma ~ dunif(0,20),
  a_shelf[shelf] ~ dnorm(0,sigma_shelf),
  sigma_shelf ~ dcauchy(0,1),
  a_acs[acs] ~ dnorm(0,sigma_acs),
  sigma_acs ~ dcauchy(0,1)),
  data3.shelf,
  chains = 4)

precis(shelf.stan,depth=2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(shelf.stan))


```

## Problem 2: Smoking deaths among doctors

In 1961 Doll and Hill sent out a questionnaire to all men on the British Medical Register inquiring about their smoking habits. Almost 70% of such men replied. Death certificates were obtained for medical practitioners and causes of death were assigned on the basis of these certificates. The breslow data set contains the person-years of observations and deaths from coronary artery disease accumulated during the first ten years of the study.

Analyse this data set to determine the posterior probability that smoking increases death by coronary artery disease, that age increases death by coronary artery disease, and that there is an interaction between age and smoking.

You can load the data set and learn about the columns using the commands below

```{r}
data("breslow",package = "boot")
help("breslow",package ="boot")

d <- breslow

d$age <- as.integer(d$age)
d$n <- as.integer(d$n)

str(d)
summary(d)

#fit a model with smoke
m.smoke <- map2stan(
 alist(
  y ~ dbinom( n , p ),
  logit(p) <- a + b_s*smoke,
  a ~ dnorm(0, 5),
  b_s ~ dnorm(0, 2)),
 data=d , iter=4000,chains=4 )

precis(m.smoke)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(m.smoke))

logistic(-5.96)#the posterior probability that smoking increases death by coronary artery disease is 0.0025?

#fit a model with age
m.age <- map2stan(
 alist(
  y ~ dbinom( n , p ),
  logit(p) <- a_age[age],
  a_age[age] ~ dnorm(a, sigma),
  a ~ dnorm(0, 5),
  sigma ~ dcauchy(0, 1)
 ), data=d , iter=4000 , chains=4 )

precis(m.age,depth = 2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(m.age,depth = 2))

logistic(-7.63)
logistic(-6.14)
logistic(-4.99)
logistic(-3.92)#more older more possible to dead by coronary artery disease?

#fit a model with age and smoke
m.smoke.age <- map2stan(
 alist(
  y ~ dbinom( n , p ),
  logit(p) <- a_age[age] + b_s*smoke,
  a_age[age] ~ dnorm(a, sigma),
  a ~ dnorm(0, 5),
  sigma ~ dcauchy(0, 1),
  b_s ~ dnorm(0, 2)), 
 data=d , iter=4000 , chains=4 )

precis(m.smoke.age,depth=2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(m.smoke.age,depth=2))

#add interaction between age and smoke
m.smoke.age.interaction <- map2stan(
 alist(
  y ~ dbinom( n , p ),
  logit(p) <- a_age[age] + b_s*smoke + b_as[age]*smoke,
  a_age[age] ~ dnorm(a, sigma),
  a ~ dnorm(0, 5),
  sigma ~ dcauchy(0, 1),
  c(b_s,b_as) ~ dnorm(0, 2)), 
 data=d , iter=4000 , chains=4 )

precis(m.smoke.age.interaction,depth=2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(m.smoke.age.interaction,depth=2))
```

You can think of "person-years" as the number of observations

Note: You almost certainly have the `boot` package on your computer, but if you do not have the `boot` package on your computer then you will need to `install.packages("boot")`

Note: do NOT do `library(boot)`.  This will make the `logit` function from boot over-ride the one from `rethinking`

## Problem 3: Cane Sugar

This data comes from an experiment to measure disease resistance in different varieties of sugar cane.

Is there evidence of differences in disease resistance in the different varieties?  Does including an adaptive prior for Block improve the model fit?

You can get the data and learn about it with:

```{r}
data("cane",package="boot")
help("cane",package="boot")
head(cane)
summary(cane)
str(cane)

cane$block2 <- coerce_index(cane$block)
cane$n <- as.integer(cane$n)
cane$var2 <- coerce_index(cane$var)
cane$plot <- 1:nrow(cane)
cane$pro <- cane$r/cane$n

#fit a model with different varieties
m.var <- map2stan(
 alist(
  r ~ dbinom( n , p ),
  logit(p) <- a + a_var[var2],
  a ~ dnorm(0,2),
  a_var[var2] ~ dnorm(0, 5)),
 data=cane, iter=4000, chains=4)

precis(m.var,depth=2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(m.var,depth=2))

#fit a model with different varieties and a adaptive prior for Block
m.var.block <- map2stan(
 alist(
  r ~ dbinom( n , p ),
  logit(p) <- a + a_var[var2] + a_block[block2],
  a ~ dnorm(0,2),
  a_var[var2] ~ dnorm(0, 5),
  a_block[block2] ~ dnorm(0, sigma),
  sigma ~ dcauchy(0,1)),
 data=cane, iter=4000, chains=4)

precis(m.var.block,depth=2)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(m.var.block,depth=2))

#compare the models
compare(m.var,m.var.block)
```


