---
title: "Chapter_14"
author: "Xiaoyan Yue"
date: "2/22/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##14E1  
Ti ~ Poisson(mu)
log(mu)= a + b*log(P_est)
log(P_obs) ~ Normal (log(P_est),P_SD)
a ~ Normal(0,10)
b ~ Normal(0,1)


##14M3
###First, repeat model in the chapter
```{r}
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

dlist <- list(
  div_obs=d$Divorce,
  div_sd=d$Divorce.SE,
  R=d$Marriage,
  A=d$MedianAgeMarriage
)

m14.1 <- map2stan(
  alist(
    div_est ~ dnorm(mu,sigma),
    mu <- a + bA*A + bR*R,
    div_obs ~ dnorm(div_est,div_sd),
    a ~ dnorm(0,10),
    bA ~ dnorm(0,10),
    bR ~ dnorm(0,10),
    sigma ~ dcauchy(0,2.5)
  ) ,
  data=dlist ,
  start=list(div_est=dlist$div_obs) ,
  WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
  control=list(adapt_delta=0.95) )

precis(m14.1,depth = 2)
```

###Second, double the standard error
```{r}
data(WaffleDivorce)
d <- WaffleDivorce

d$Divorce.SE_double <- 2*d$Divorce.SE
head(d)

dlist <- list(
  div_obs=d$Divorce,
  div_sd=d$Divorce.SE_double,
  R=d$Marriage,
  A=d$MedianAgeMarriage
)

m14.M3 <- map2stan(
  alist(
    div_est ~ dnorm(mu,sigma),
    mu <- a + bA*A + bR*R,
    div_obs ~ dnorm(div_est,div_sd),
    a ~ dnorm(0,10),
    bA ~ dnorm(0,10),
    bR ~ dnorm(0,10),
    sigma ~ dcauchy(0,2.5)
  ) ,
  data=dlist ,
  start=list(div_est=dlist$div_obs) ,
  WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2,
  control=list(adapt_delta=0.95))

precis(m14.M3,depth = 2)
```
  
###There is no many differences in the posterior of bA and bR between two models, but increasing the standard error of the outcome do decrease the posterior of the sigma in the model.


##14H1  
###Load the data
```{r}
data("elephants")
d1 <- elephants
#look at the data
head(d1)
#log the AGE
d1$log_age <- log(d1$AGE)
str(d1)
summary(d1)
```
  
###Fit a poisson model
```{r}
m14.H1 <- map2stan(
  alist(
    MATINGS ~ dpois( lambda ) ,
    log(lambda) <- a + bA*log_age,
    a ~ dnorm(0,10),
    bA ~ dnorm(0,1)),
  data=d1 , iter=3000 , warmup=1000 , chains=4 )

precis(m14.H1,depth=2)
```
###Fit a model with measurement errors with a standard error of +/- 5 years
```{r}
dlist <- list(
  MATINGS = d1$MATINGS,
  AGE_obs=d1$log_age,
  AGE_sd=rep(log(5),length(d1$log_age))
)

m14.H1.sd <- map2stan(
  alist(
    MATINGS ~ dpois( lambda ) ,
    log(lambda) <- a + bA*AGE_est[i],
    AGE_obs ~ dnorm(AGE_est,AGE_sd),
    a ~ dnorm(0,10),
    bA ~ dnorm(0,1)),
  data=dlist ,
  start=list(AGE_est=dlist$AGE_obs) ,
  WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2,
  control=list(adapt_delta=0.95))

precis(m14.H1.sd,depth = 2)

```
  
###After including the standard errors, the positive relationship between age and matting are going to disappear. The posterior mean of b changed from 2.02 to 0.32.  

  
##14H2
###Increase the sd value to 100, make the posterior mean to reach zero 
```{r}
dlist <- list(
  MATINGS = d1$MATINGS,
  AGE_obs=d1$log_age,
  AGE_sd=rep(log(100),length(d1$log_age))
)

m14.H1.sd.increase <- map2stan(
  alist(
    MATINGS ~ dpois( lambda ) ,
    log(lambda) <- a + bA*AGE_est[i],
    AGE_obs ~ dnorm(AGE_est,AGE_sd),
    a ~ dnorm(0,10),
    bA ~ dnorm(0,1)),
  data=dlist ,
  start=list(AGE_est=dlist$AGE_obs) ,
  WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2,
  control=list(adapt_delta=0.95))

precis(m14.H1.sd.increase,depth = 2)
```
  
###It is going to reache zero, when the standard error gets to 100.
