---
title: "Chapter_14_2"
author: "Xiaoyan Yue"
date: "3/1/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache= TRUE, autodep = TRUE)

library(rethinking)
library(reshape2)
library(ggplot2)
library(brms)
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

##14E2
Ti ~ Poisson(mu)  
log(mu)= a + b*logP  
logP ~ Normal (V_p,sigma_p)  
V_p ~ Normal(0.5,1)  
a ~ Normal(0,10)  
b ~ Normal(0,1)  
sigma_p ~ Cauchy(0,1)
  
##14M1 Using the mathematical form of the imputation model in the chapter, expalin what is being assumed about how the missing values were generated.  
####It assumed that the missing value is generated completely at random.

  
##14M2
###load the data
```{r}
data(milk)
```

###Repeat the model in Chapter 6
```{r}
#Remove NA 
d1 <- milk[ complete.cases(milk) , ]
d1$neocortex <- d1$neocortex.perc / 100
dim(d1)
a.start <- mean(d1$kcal.per.g)
sigma.start <- log(sd(d1$kcal.per.g))
#Fit different model
m6.11 <- map(
  alist(
    kcal.per.g ~ dnorm( a , exp(log.sigma) )
  ) ,
  data=d1 , start=list(a=a.start,log.sigma=sigma.start) )
m6.12 <- map(
  alist(
    kcal.per.g ~ dnorm( mu , exp(log.sigma) ) ,
    mu <- a + bn*neocortex
  ) ,
  data=d1 , start=list(a=a.start,bn=0,log.sigma=sigma.start) )
m6.13 <- map(
  alist(
    kcal.per.g ~ dnorm( mu , exp(log.sigma) ) ,
    mu <- a + bm*log(mass)
  ) ,
  data=d1 , start=list(a=a.start,bm=0,log.sigma=sigma.start) )
m6.14 <- map(
  alist(
    kcal.per.g ~ dnorm( mu , exp(log.sigma) ) ,
    mu <- a + bn*neocortex + bm*log(mass)
  ) ,
  data=d1 , start=list(a=a.start,bn=0,bm=0,log.sigma=sigma.start) )

#Compare the models
(milk.models <- compare( m6.11 , m6.12 , m6.13 , m6.14 ))

```
  
###Imputing neocortex
```{r}
#Load the data
d2 <- milk
d2$neocortex.prop <- d2$neocortex.perc / 100
d2$logmass <- log(d2$mass)
summary(d2)

#Create the data list
data_list <- list(
  kcal = d2$kcal.per.g,
  neocortex = d2$neocortex.prop,
  logmass = d2$logmass )

#Fit different models
m14M2.1 <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ),
    mu <- a,
    a ~ dnorm(0,100),
    sigma ~ dcauchy(0,1)
  ) ,
  data=data_list , iter = 1e4, chains = 2 )

m14M2.2 <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ),
    neocortex ~ dnorm(nu,sigma_N),
    mu <- a + bN*neocortex,
    a ~ dnorm(0,100),
    bN ~ dnorm(0,10),
    nu ~ dnorm(0.5,1),
    sigma_N ~ dcauchy(0,1),
    sigma ~ dcauchy(0,1)
  ) ,
  data=data_list , iter = 1e4, chains = 2 )

m14M2.3 <- map2stan(
  alist(
    kcal ~ dnorm(mu,sigma),
    mu <- a + bM*logmass,
    a ~ dnorm(0,100),
    bM ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)
  ) ,
  data=data_list , iter=1e4 , chains=2 )

m14M2.4 <- map2stan(
  alist(
    kcal ~ dnorm(mu,sigma),
    mu <- a + bN*neocortex + bM*logmass,
    neocortex ~ dnorm(nu,sigma_N),
    a ~ dnorm(0,100),
    c(bN,bM) ~ dnorm(0,10),
    nu ~ dnorm(0.5,1),
    sigma_N ~ dcauchy(0,1),
    sigma ~ dcauchy(0,1)
  ) ,
  data=data_list , iter=1e4 , chains=2 )
#Compare the models
(milk.models1 <- compare( m14M2.1 , m14M2.2 , m14M2.3 , m14M2.4 ))

```
  
####It seems the weight of the model with imputation increased to 0.95 from 0.91 in the model excluded the missing data, and the standared error also decreased 
  
##14H3
###create the data
```{r}
set.seed(100)
x <- c(rnorm(10), NA)
y <- c(rnorm(10,x),100)
d <- list(x=x,y=y)
```

###Fit the model by imputing the one missing value for x
```{r}
m14H3.2 <- map2stan(
  alist(
    y ~ dnorm(mu,sigma),
    mu <- a + bX*x,
    x ~ dnorm(0,1),
    a ~ dnorm(0,100),
    bX ~ dnorm(0,100),
    sigma ~ dcauchy(0,1)
  ) ,
  data=d , iter=1e4 , chains=2 )

precis(m14H3.2)
```
####I think now it is not a strong positive relationship between two variables.The mean posterior of bX is 7.08, and the standared error is 25.74, almost half of the values are lower than 0. Because we defined a certain normal distribution for the missing value in x, it makes sampling bias. I think may be we need to specify the distribution of the missing value xi based on the relationships between x and y. 

