---
title: "Assignment_Chapter_12"
author: "XIAOYAN YUE"
date: "2016Äê10ÔÂ27ÈÕ"
output: 
  html_document: 
    keep_md: yes
---

##12E1 
As smaller sample sizes, shrink more. Hence, (a)will create more shrinkage.

##12E2
yi ~ Binomial(1,pi)
logit(pi) = a_group[i]+bxi
a_group ~ Normal(a,sigma) #creat varing intercepts prior
b ~ Normal(0,1)
a ~ Normal(0,1)#prior for average group
sigma ~ HalfCauchy(0,1)#prior for standard deviation of groups

The prior for the a_group intercept is now a function of two parameters,a and sigma. The Gaussian distribution with mean 'a' and standard deviation 'sigma' is the prior for each tank's intercept. Hence, there are two levels in the model,each resembling a simpler model.

##12M1
###Revisit the Reed frog survival data, data(reedfrogs), and add the predation and size treatment variables to the varying intercepts model. Consider models with either main effect alone, both main effects, as well as a model including both and their interaction. Instead of focusing on inferences about these two predictor variables, focus on the inferred variation across tanks. Explain why it changes as it does across models.
```{r}
library(rethinking)

data(reedfrogs)
d <- reedfrogs
head(d)
colnames(d)
d
# make the tank cluster variable
d$tank <- 1:nrow(d)

#make dummy variables for predation and size
d$pred2 <- ifelse(d$pred=="no",0,1)
d$size2 <- ifelse(d$size=="small",0,1)

#fit the model with either main effect alone
##pred
m12m.1.pred <- map2stan(
 alist(
  surv ~ dbinom( density , p ) ,
  logit(p) <- a_tank[tank] + bp*pred2,
  a_tank[tank] ~ dnorm( a , sigma ) ,
  bp ~ dnorm(0,2),
  a ~ dnorm(0,5) ,
  sigma ~ dcauchy(0,1)
 ), data=d , iter=4000 , chains=4 )

precis(m12m.1.pred,depth=2)

##size
m12m.1.size <- map2stan(
 alist(
  surv ~ dbinom( density , p ) ,
  logit(p) <- a_tank[tank] + bs*size2,
  a_tank[tank] ~ dnorm( a , sigma ) ,
  bs ~ dnorm(0,2),
  a ~ dnorm(0,5) ,
  sigma ~ dcauchy(0,1)), 
 data=d , iter=4000 , chains=4 )

precis(m12m.1.size,depth=2)

##fit models with both main effects, including pred and size
m12m.1.pred.size <- map2stan(
 alist(
  surv ~ dbinom( density , p ) ,
  logit(p) <- a_tank[tank] + bp*pred2 + bs*size2 ,
  a_tank[tank] ~ dnorm( a , sigma ) ,
  bp ~ dnorm(0,2),
  bs ~ dnorm(0,2),
  a ~ dnorm(0,5) ,
  sigma ~ dcauchy(0,1)), 
 data=d , iter=4000 , chains=4 )

precis(m12m.1.pred.size,depth=2)

##fit model including both and their interaction
m12m.1.pred.size.interaction <- map2stan(
 alist(
  surv ~ dbinom( density , p ) ,
  logit(p) <- a_tank[tank] + bp*pred2 + bs*size2 +bps*pred2*size2,
  a_tank[tank] ~ dnorm( a , sigma ) ,
  bp ~ dnorm(0,2),
  bs ~ dnorm(0,2),
  bps ~ dnorm(0,1),
  a ~ dnorm(0,5) ,
  sigma ~ dcauchy(0,1)), 
 data=d , iter=4000 , chains=4 )

precis(m12m.1.pred.size.interaction,depth=2)

```

##12M2
### Compare the models you fit just above, using WAIC. Can you reconcile the differences in WAIC with the posterior distributions of the models?
```{r}
#compare the models
compare(m12m.1.pred,m12m.1.size,m12m.1.pred.size,m12m.1.pred.size.interaction)
```

#The function:link, which is trying to get more predictions
#extract.sample: which only get samples from the real posterior distributions

##12M3 Re-estimate the basic Reed frog varying intercept model, but now using a Cauchy distribution in place of the Gaussian distribution for the varying intercepts. That is, fit this model: si ~ Binomial(ni; pi); logit(pi) = a_tank[i]; a[tank] ~ Cauchy(a,sigma); a ~ Normal(0; 1); sigma ~ HalfCauchy(0; 1). Compare the posterior means of the intercepts, a[tank], to the posterior means produced in the chapter, using the customary Gaussian prior. Can you explain the pattern of differences?

```{r}
library(rethinking)
data(reedfrogs)
d <- reedfrogs
d$tank <- 1:nrow(d)
head(d)
str(d)

d$pred <- as.numeric(d$pred)
d$size <- as.numeric(d$size)


#fit the intercepts with a Gaussian distribution
m12m.3a <- map2stan(
 alist(
  surv ~ dbinom( density, p ) ,
  logit(p) <- a_tank[tank],
  a_tank[tank] ~ dnorm( a , sigma ) ,
  a ~ dnorm(0,1) ,
  sigma ~ dcauchy(0,1)), 
 data=d , iter=4000 , chains=4 )

precis(m12m.3a,depth=2)

#fit intercepts with a Cauchy distribution
m12m.3b <- map2stan(
 alist(
  surv ~ dbinom( density, p ) ,
  logit(p) <- a_tank[tank],
  a_tank[tank] ~ dcauchy( a , sigma ) ,
  a ~ dnorm(0,1) ,
  sigma ~ dcauchy(0,1)), 
 data=d , iter=4000 , chains=4 )
 
precis(m12m.3b,depth=2)

#it seems the intercept with a cauchy distribution displayed a higher variation (from -2 to 12) than that with a normal distribution (from -1.82 to 3.9). If it the priors of the intercepts with cauchy distribution created a higher variation among different ponds...?

```

##12H1
```{r}
#get the data
data(bangladesh)
d2 <- bangladesh 

str(d2)

#To ensure that the cluster variable is a contiguous set of integers
sort(unique(d2$district))

#make a new variable that is contiguous
d2$district_id <- as.integer(as.factor(d2$district))
sort(unique(d2$district_id))

#fit a traditional fixed-effects model that uses dummy variable for district
colnames(d2) <- sub(".","_",fix= TRUE,colnames(d2))

str(d2)
summary(d2)

m12h.1.traditional <- map2stan(
 alist(
  use_contraception ~ dbinom( 1 , p ) ,
  logit(p) <- a_district_id [district_id], 
  a_district_id[district_id] ~ dnorm( 0, 1 )), 
 data=d2 , iter=4000 , chains=4 )

precis(m12h.1.traditional,depth=2)
plot(m12h.1.traditional,ask=FALSE)

#fit a mutilevel model with varying intercepts for districts
m12h.1.mutilevel <- map2stan(
 alist(
  use_contraception ~ dbinom( 1 , p ) ,
  logit(p) <- a_district_id[district_id],
  a_district_id[district_id] ~ dnorm( a , sigma ) ,
  a ~ dnorm(0,1) ,
  sigma ~ dcauchy(0,1)
 ), data=d2 , iter=4000 , chains=4 )

precis(m12h.1.mutilevel,depth=2)

#plot the predicted proportions of women in each district using contraception for traditional model
#extract posterior 
library(reshape2)

pred.df <- data.frame(district_id=unique(d2$district_id))
link.vary <- link(m12h.1.traditional,data=pred.df,n=4000)

pred.df$trad.vary.link <- apply(link.vary,2,mean)
pred.df$trad.vary.coef <- logistic(coef(m12h.1.traditional)[1:60])

post_traditional.vary <- extract.samples(m12h.1.traditional)$a_district
post_mutilevel.vary <- extract.samples(m12h.1.mutilevel)$a_district


```
DCE. 2
##12E3 Make the following model into a mutilevel model
yi ~ Normal(mu, sigma)  
mu = a + a_group[group] + bxi  
a_group[group] ~ Normal(0,sigma_group)  
a ~ Normal(0,10)  
b ~ Normal(0,1)  
sigma ~ HalfCauchy(0,2)  
sigma_group ~ Normal(0,1)  

##12E4 Write an example mathmatical model formula for a Poisson regression with varying intercepts.
Based on the example from Chapter 10, T is total_tools, P is population, and i indexes each society:  
Ti ~ Poisson(mu)  
log(mu) = a + a_society[society] + bplogPi  
a ~ Normal(0,10)  
bp ~ Normal(0,1)  
a_society[society] ~ Normal(0,sigma_society)  
sigma_society ~ HalfCauchy(0,1)  

##12E5 Write an example mathematical model formula for a Poisson regression with two different kinds of varying intercepts, a cross-classified model.
Based on the former example, I added a variation as "region" for different societies:  
Ti ~ Poisson(mu)  
log(mu) = a + a_society[society] + a_region[region] + bplogPi  
a ~ Normal(0,10)  
bp ~ Normal(0,1)  
a_society[society] ~ Normal(0,sigma_society)  
sigma_society ~ HalfCauchy(0,1)  
a_region[region] ~ Normal(0, sigma_region)  
sigma_region ~ HalfCauchy(0,1)  

##12M4 Fit the following cross-classified multilevel model to the chimpanzees data:

Li ~ Binomial(1, pi)  
logit(pi) = a_actor[i] + a_block[i] + (bP + bPCCi)Pi  
a_actor ~ Normal(a, sigma_actor)  
a_block ~ Normal(gama, a_block)   
a,gama,bP,bPC ~ Normal(0, 10)  
a_actor, a_block ~ HalfCauchy(0, 1)  

```{r}
#The model in the Chapter
library(rethinking)
data(chimpanzees)
d2 <- chimpanzees
d2$recipient <- NULL     # get rid of NAs
d2$block_id <- d2$block  # name 'block' is reserved by Stan

m12m4.1 <- map2stan(
  alist(
    pulled_left ~ dbinom( 1 , p ),
    logit(p) <- a + a_actor[actor] + a_block[block_id] +
      (bp + bpc*condition)*prosoc_left,
    a_actor[actor] ~ dnorm( 0 , sigma_actor ),
    a_block[block_id] ~ dnorm( 0 , sigma_block ),
    c(a,bp,bpc) ~ dnorm(0,10),
    sigma_actor ~ dcauchy(0,1),
    sigma_block ~ dcauchy(0,1)
  ) ,
  data=d2, warmup=1000 , iter=6000 , chains=4 , cores=3 )
#fit the cross-classfied multilevel model to the chimpanzees data (without the common intercept for two types of clusters)
m12m4.2 <- map2stan(
  alist(
    pulled_left ~ dbinom( 1 , p ),
    logit(p) <- a_actor[actor] + a_block[block_id] +
      (bp + bpc*condition)*prosoc_left,
    a_actor[actor] ~ dnorm( a , sigma_actor ),
    a_block[block_id] ~ dnorm( gama , sigma_block ),
    c(a,gama,bp,bpc) ~ dnorm(0,10),
    sigma_actor ~ dcauchy(0,1),
    sigma_block ~ dcauchy(0,1)
  ) ,
  data=d2, warmup=1000 , iter=6000 , chains=4 , cores=3 )

#compare two different model
precis(m12m4.1,depth = 2)
precis(m12m4.2,depth = 2)

compare(m12m4.1,m12m4.2)

#extract samples from two different model and plot
pairs(m12m4.1)
paris(m12m4.1)
```
