---
title: "Assignment_Chapter_09"
author: "XIAOYAN YUE"
data: "1st/August/2016"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---

```{r}
##load the data and libraries
setwd("D:/")
data <- read.csv("TomatoR2CSHL.csv")

library(rethinking)
library(ggplot2)
library(reshape2)

#take a look at the data TO figure out if the data are normal distribution 
head(data)
summary(data)

#denSity plots
pl <- ggplot(data = data, aes(x=hyp,fill=trt))
pl <- pl + geom_density()
pl <- pl + facet_grid(species ~ trt)
pl

#box plots
pl <- ggplot(data = data, aes(y=hyp,x=species,fill=trt))
pl <- pl + geom_boxplot()
pl

##data transformation (if the original data are not so normal distribution)

#Square root transformation
pl <- ggplot(data = data, aes(x=sqrt(hyp),fill=trt))
pl <- pl + geom_density()
pl <- pl + facet_grid(species ~ trt)
pl + ggtitle("sqrt transformed")

#Log2 transformation
pl <- ggplot(data = data, aes(x=log2(hyp),fill=trt))
pl <- pl + geom_density()
pl <- pl + facet_grid(species ~ trt)
pl + ggtitle("log2 transformed")


##except by plotting the data to see if it is normal distribution, we can use normality test by shapiro.test
by(data$hyp,list(data$trt,data$species),shapiro.test)
by(sqrt(data$hyp),list(data$trt,data$species),shapiro.test)
by(log2(data$hyp),list(data$trt,data$species),shapiro.test)
?by #function 'by' is an object-oriented wrapper for tapply applied to data frame

#so, squrt transformed is the best

#categorical variable for trt
data$trtL <- ifelse(data$trt=="L",1,0)

#alternative way
levels(data$trt)
data$trt2 <- as.numeric(data$trt)-1 #0 = H, 1=L

#################################################
###what I did for the variables
#set the dummy variable for the data
##trt
#data$trt <- ifelse(dat1$trt=="H",1,0)
#summary(dat1)
##species
#data$species_chilense <- ifelse(dat1$species=="S. chilense",1,0)
#data$species_chimielewskii <- ifelse(dat1$species=="S. chimielewskii",1,0)
#data$species_habrochaites <- ifelse(dat1$species=="S. habrochaites",1,0)
#data$species_pennellii <- ifelse(dat1$species=="S. pennellii",1,0)
#data$species_peruvianum <- ifelse(dat1$species=="S. peruvianum",1,0)
###what I did for the variables
####################################################

#categorical variables for species
#almost the same thing as the last step, set the dummy variable "species" into different column, which is really useful when you have a lot of species, you can not do it one by one
data$index <- 1:nrow(data)
data2 <- dcast(data,index + hyp + trt2 ~ species, value.var = "species",fun.aggregate = length)

head(data2)

#set a model with a prior distribution using map2stan
##stan needs a "clear data", we must subset the data frame to contain only the relevant columns and get rid of the other variables
data2.trt <- data2[, c("hyp","trt2")]
head(data2.trt)

#set the model for hyp and trt
#trt model
trt.stan <- map2stan(alist(
  hyp ~ dnorm(mu,sigma),
  mu <- a + bT * trt2,
  a ~ dnorm(0,100),
  bT ~ dnorm(0,10),
  sigma ~ dunif(0,20)),
  data2.trt,
  chains = 4)

#test the model
plot(trt.stan)
precis(trt.stan)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(trt.stan))


##set the model for hyp and species
#first, fix species names to get rid of space 
colnames(data2) <- sub(". ","_",colnames(data2))
head(data2)

#model where each species has its own intercept
data2.species.all <- data2[,c(2,4:8)]
head(data2.species.all)

#set the model for hyp and species
species.stan <- map2stan(alist(
  hyp ~ dnorm(mu, sigma),
  mu <- bChil*S_chilense + bChmi*S_chmielewskii + bHab*S_habrochaites + bPen*S_pennellii + bPer*S_peruvianum,
  c(bChil, bChmi, bHab, bPen, bPer) ~ dnorm(33.35, 20),
  sigma ~ dunif(0,20)),
  data2.species.all,
  chains = 4)

#test the model
plot(species.stan)
precis(species.stan)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(species.stan))

#use S. chmi as intercept
data2.species.intercept <- data2[,c(2,4,6:8)]

head(data2.species.intercept)


species.stan.intercept <- map2stan(alist(
  hyp ~ dnorm(mu,sigma),
  mu <- a + bChil*S_chilense + bHab*S_habrochaites + bPen * S_pennellii + bPer*S_peruvianum,
  a ~ dnorm(33.35,10),
  c(bChil,bHab,bPen,bPer) ~ dnorm(0,10),
  sigma ~ dunif(0,20)),
  data2.species.intercept,
  chains = 4)

#test the model
plot(species.stan.intercept)
precis(species.stan.intercept)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(species.stan.intercept))

#set the model with species and trt, except the intercept

data2.species.trt <- data2[,c(2:4,6:8)]

head(data2.species.trt)

species.trt.stan <- map2stan(alist(
  hyp ~ dnorm(mu,sigma),
  mu <- a + bT*trt2 + bChil*S_chilense + bHab*S_habrochaites + bPen * S_pennellii + bPer*S_peruvianum,
  a ~ dnorm(33.35,10),
  c(bT,bChil,bHab,bPen,bPer) ~ dnorm(0,10),
  sigma ~ dunif(0,20)),
  data2.species.trt,
  chains = 4)

plot(species.trt.stan)
precis(species.trt.stan)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(species.trt.stan))

#compare the models

compare(trt.stan,species.trt.stan,species.stan.intercept,species.stan)

plot(compare(trt.stan,species.trt.stan,species.stan.intercept,species.stan))


coeftab(trt.stan,species.trt.stan,species.stan.intercept,species.stan)

plot(coeftab(trt.stan,species.trt.stan,species.stan.intercept,species.stan))

#what can we say about the treatment effect?

#get the posterior distribution of bT, the treatment coefficient.

post.bT <- extract.samples(species.trt.stan)$bT

dens(post.bT,show.HPDI = 0.95)
#the fact that the 95% HDPI intervals are far away from 0 is strong evidence that bT is positive

# what percent of the posterior distribution of bT is less than or equal to ?
sum(post.bT <= 0) / length(post.bT) 
# turns out None of the posterior distribution for bT is less than or equal to 0.
```
##2a) Redo the tomato hyp analysis (or at least the final model with species and treatment) but using square root transformation.  Are model fits improved?  Does it change the parameter estimates?  
```{r}
#set the square root transformation
data2$hyp_sqrt <- sqrt(data2$hyp)

head(data2)

#set the model with species and trt, except the intercept

data2_sqrt <- data2[,c(3:4,6:9)]

head(data2_sqrt) 

mean(data2_sqrt$hyp_sqrt)

data2_sqrt.stan <- map2stan(alist(
  hyp_sqrt ~ dnorm(mu,sigma),
  mu <- a + bT*trt2 + bChil*S_chilense + bHab*S_habrochaites + bPen * S_pennellii + bPer*S_peruvianum,
  a ~ dnorm(5.7,10),
  c(bT,bChil,bHab,bPen,bPer) ~ dnorm(0,10),
  sigma ~ dunif(0,20)),
  data2_sqrt,
  chains = 4)

plot(data2_sqrt.stan)
precis(data2_sqrt.stan)
par(mfrow=c(1,1),mfcol=c(1,1))
plot(precis(data2_sqrt.stan))

#compare the model using square root transformation and the original data
compare(data2_sqrt.stan,species.trt.stan)
plot(compare(data2_sqrt.stan,species.trt.stan))


coeftab(data2_sqrt.stan,species.trt.stan)
plot(coeftab(data2_sqrt.stan,species.trt.stan))

```

##2b) Is WAIC an appropriate way to compare the model fits with untransformed and transformed data?  How can you test this?  
For the results of compare, we see the model fits with tranformed data has a much lower WAIC and the largest weight value.  
  
How to test?

##2c) How do you get sqrt transformed parameter estimates (especially when they are < 1) back on the original scale (I don't know the answer to this)  


##3)  Fit a model with a species by treatment interaction.  Which species have a trt response that is different from S. chm?  

```{r}
head(data2)

#Fit a model with a species by treatment interaction.
#S. chm?





```



