---
title: "Assignment_Chapter_07"
author: "XIAOYAN YUE"
date: "06/04/2016"
output: 
  html_document: 
    keep_md: yes
---
###7E1. For each of the causal relationships below, name a hypothetical third variable that would lead to an interaction effect.

####(1) Bread dough rises because of yeast.
time, temperature

####(2) Education leads to higher income.
family background,country

####(3) Gasoline makes a car go.
physical situation of the car

###7E2. Which of the following explanations invokes an interaction?
\(2)

###7E3. For each of the explanations in 7E2, write a linear model that expresses the stated relationship.
(1)caramelized ~ Normal(mu,sigma)  
mu = a + b.h\*heat + b.w\*water

(2)car_faster ~ Normal(mu,sigma)  
mu = a + b_c\*cylinder + b_f\*fuel_injector + b_cf\*cylinder\*fuel_injector

(3)politicol_beliefs ~ Normal(mu,sigma)  
mu = a + b_p\*parent_beliefs + b_f\*friends_beliefs (because their friends' beliefs didnot effect their parents' beliefs)

(4)intelligent_animal ~ Normal(mu,sigma)  
mu = a + b_s\*social+b_a*appendages

###7M1
Under hot temperature, at any water and light conditionsno tupips could grow blooms. 

###7M2
temperature <- c(0,1)  
regression equiation:
bloom_size ~ water:shade:cool + water:cool + shade:cool

###7M3
```{r}
##invent a set of data? not just creat a model?
```

###7H1
```{r}
library(rethinking)
data(tulips)
d <- tulips

##make centered verison of shade and water
d$shade.c <- d$shade - mean(d$shade)
d$water.c <- d$water - mean(d$water)

#set the dummy variable for the data
d$bed_a <- ifelse(d$bed=="a",1,0)
d$bed_b <- ifelse(d$bed=="b",1,0)
d$bed_c <- ifelse(d$bed=="c",1,0)

##creat the model
m.7H1 <- map(
  alist(
    blooms ~ dnorm( mu , sigma ) ,
    mu <- a + bW*water.c + bS*shade.c + bWS*water.c*shade.c + bA*bed_a + bB*bed_b + bC*bed_c, 
    a ~ dnorm( 130 , 100 ) ,
    bW ~ dnorm( 0 , 100 ) ,
    bS ~ dnorm( 0 , 100 ) ,
    bWS ~ dnorm( 0 , 100 ) ,
    bA ~ dnorm( 0, 1 ),
    bB ~ dnorm( 0, 1 ),
    bC ~ dnorm( 0, 1),#set the prior for the added parameter
    sigma ~ dunif( 0 , 100 )
  ) ,
  data=d ,
  start=list(a=mean(d$blooms),bW=0,bS=0,bWS=0,bA=0,bB=0,bC=0,sigma=sd(d$blooms)) )
precis(m.7H1)
plot(precis(m.7H1))
```

####7H2
```{r}
#fit the model omited bed
m.7H2 <- map(
  alist(
    blooms ~ dnorm( mu , sigma ) ,
    mu <- a + bW*water.c + bS*shade.c + bWS*water.c*shade.c ,
    a ~ dnorm( 130 , 100 ) ,
    bW ~ dnorm( 0 , 100 ) ,
    bS ~ dnorm( 0 , 100 ) ,
    bWS ~ dnorm( 0 , 100 ) ,
    sigma ~ dunif( 0 , 100 )
  ) ,
  data=d ,
  start=list(a=mean(d$blooms),bW=0,bS=0,bWS=0,sigma=sd(d$blooms)) )

compare(m.7H1,m.7H2)#compare these two models using WAIC
```
It seems the bed variable did not effect the outcomes in this study.

```{r}
#what means reconcile the WAIC results?
```

####7H3
```{r}
library(rethinking)
data(rugged)
d <- rugged

# make log version of outcome
d$log_gdp <- log( d$rgdppc_2000 )

# extract countries with GDP data
dd <- d[ complete.cases(d$rgdppc_2000) , ]

#(a) using map to fit just the interaction model without country Seychelles

# extract country Seychelles
dd1 <- dd[dd$country!="Seychelles",]
#fit the model
m.7H3_1 <- map(
  alist(
    log_gdp ~ dnorm( mu , sigma ) ,
    mu <- a + bR*rugged + bAR*rugged*cont_africa + bA*cont_africa,
    a ~ dnorm( 8 , 100 ) ,
    bA ~ dnorm( 0 , 1 ) ,
    bR ~ dnorm( 0 , 1 ) ,
    bAR ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=dd1 )

#using the same model fit to the full data
m.7H3_2 <- map(
  alist(
    log_gdp ~ dnorm( mu , sigma ) ,
    mu <- a + bR*rugged + bAR*rugged*cont_africa + bA*cont_africa,
    a ~ dnorm( 8 , 100 ) ,
    bA ~ dnorm( 0 , 1 ) ,
    bR ~ dnorm( 0 , 1 ) ,
    bAR ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=dd )
#compare the inference of two models
precis(m.7H3_1)
precis(m.7H3_2)

#it seems the only difference is the interaction effect is smaller in the model without country Seychelles
```

(b)plot the posterior predictions for above two models
```{r}
#set the value of rugged
rugged.seq <- seq(from=-1,to=8,by=0.25)

# compute mu over samples from model m.7H3_1 (without country Seychelles)
mu.NotAfrica <- link( m.7H3_1 , data=data.frame(cont_africa=0,rugged=rugged.seq) )#based on model m.7H3_1 calculating mu 
mu.Africa <- link( m.7H3_1 , data=data.frame(cont_africa=1,rugged=rugged.seq) )

# summarize to means and intervals
mu.NotAfrica.mean <- apply( mu.NotAfrica , 2 , mean )
mu.NotAfrica.PI <- apply( mu.NotAfrica , 2 , PI , prob=0.97 )
mu.Africa.mean <- apply( mu.Africa , 2 , mean )
mu.Africa.PI <- apply( mu.Africa , 2 , PI , prob=0.97 )

#plot the posterior predictions for the model m.7H3_1 for non-africa country
d.A0 <- dd[dd$cont_africa==0,]
plot(log_gdp ~ rugged, data=d.A0,col=rangi2,ylab="log GDP 2000",xlab="Terrain Ruggedness Index")
mtext("Non-Africa nation variations in model m.7H3_1")
lines(rugged.seq,mu.NotAfrica.mean,col=rangi2)
shade(mu.NotAfrica.PI,rugged.seq,col = col.alpha(rangi2,0.3))

#plot the posterior predictions for The model m.7H3_1 for Africa country
d.A1 <- dd[dd$cont_africa==1,]
plot(log_gdp ~ rugged, data=d.A1,col=rangi2,ylab="log GDP 2000",xlab="Terrain Ruggedness Index")
mtext("Africa nations variations in model m.7H3_1")
lines(rugged.seq,mu.Africa.mean,col=rangi2)
shade(mu.Africa.PI,rugged.seq,col = col.alpha(rangi2,0.3))

###################
# compute mu over samples from model m.7H3_2 (full data)
mu.NotAfrica <- link( m.7H3_2 , data=data.frame(cont_africa=0,rugged=rugged.seq) )#based on model m.7H3_1 calculating mu 
mu.Africa <- link( m.7H3_2 , data=data.frame(cont_africa=1,rugged=rugged.seq) )

# summarize to means and intervals
mu.NotAfrica.mean <- apply( mu.NotAfrica , 2 , mean )
mu.NotAfrica.PI <- apply( mu.NotAfrica , 2 , PI , prob=0.97 )
mu.Africa.mean <- apply( mu.Africa , 2 , mean )
mu.Africa.PI <- apply( mu.Africa , 2 , PI , prob=0.97 )

#plot the posterior predictions for the model m.7H3_2 for Non-Africa country
d.A0 <- dd[dd$cont_africa==0,]
plot(log_gdp ~ rugged, data=d.A0,col=rangi2,ylab="log GDP 2000",xlab="Terrain Ruggedness Index")
mtext("Non-Africa nation variations in model m.7H3_2")
lines(rugged.seq,mu.NotAfrica.mean,col=rangi2)
shade(mu.NotAfrica.PI,rugged.seq,col = col.alpha(rangi2,0.3))

#plot the posterior predictions for THE model m.7H3_2 for Africa country
d.A1 <- dd[dd$cont_africa==1,]
plot(log_gdp ~ rugged, data=d.A1,col=rangi2,ylab="log GDP 2000",xlab="Terrain Ruggedness Index")
mtext("Africa nations variations in model m.7H3_1")
lines(rugged.seq,mu.Africa.mean,col=rangi2)
shade(mu.Africa.PI,rugged.seq,col = col.alpha(rangi2,0.3))
```
(c)Fit three models to the data without Seychelles and conduct a model comparison analysis using WAIC.
```{r}
# extract country Seychelles
dd1 <- dd[dd$country!="Seychelles",]

#model 1:y ~ Normal(mu,sigma)
       #mu = a + bR*R

#fit the model 1
m.c1 <- map(
  alist(
    log_gdp ~ dnorm( mu , sigma ) ,
    mu <- a + bR*rugged,
    a ~ dnorm( 8 , 100 ) ,
    bR ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=dd1 )

#model 2:y ~ Normal(mu,sigma)
       #mu = a + bA*A + bR*R 

#fit the model 2
m.c2 <- map(
  alist(
    log_gdp ~ dnorm( mu , sigma ) ,
    mu <- a + bR*rugged + bA*cont_africa,
    a ~ dnorm( 8 , 100 ) ,
    bA ~ dnorm( 0 , 1 ) ,
    bR ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=dd1 )

#model 3:y ~ Normal(mu,sigma)
       #mu = a + bA*A + bR*R + bAR*A*R 

#fit the model 3
m.c3 <- map (
  alist(
    log_gdp ~ dnorm( mu , sigma ) ,
    mu <- a + bR*rugged + bAR*rugged*cont_africa + bA*cont_africa,
    a ~ dnorm( 8 , 100 ) ,
    bA ~ dnorm( 0 , 1 ) ,
    bR ~ dnorm( 0 , 1 ) ,
    bAR ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=dd1 )

compare(m.c1,m.c2,m.c3)

##plot the model-averaged predictions of this model set

#(1)averaged-model for Non-Africa
rugged.seq <- seq(from=-1,to=8,by=0.25)
rugged.ensemble1 <- ensemble(m.c1,m.c2,m.c3,data=data.frame(cont_africa=0,rugged=rugged.seq))

# compute mu over samples from averaged-model for Non-Africa country
mu.NotAfrica <- apply (rugged.ensemble1$link, 2, mean )#based on averaged-model calculating mu 

# summarize to intervals
mu.NotAfrica.PI <- apply( rugged.ensemble1$link , 2 , PI , prob=0.97 )

#plot the posterior predictions for THE averaged-model for non-africa country
d.A0 <- dd[dd$cont_africa==0,]
plot(log_gdp ~ rugged, data=d.A0,col=rangi2,ylab="log GDP 2000",xlab="Terrain Ruggedness Index")
mtext("Model-averaged predictions for Non-Africa countrys ")
lines(rugged.seq,mu.NotAfrica,col=rangi2)
shade(mu.NotAfrica.PI,rugged.seq,col = col.alpha(rangi2,0.3))

#(2)averaged-model for Africa country
rugged.seq <- seq(from=-1,to=8,by=0.25)
rugged.ensemble2 <- ensemble(m.c1,m.c2,m.c3,data=data.frame(cont_africa=1,rugged=rugged.seq))

# compute mu over samples from averaged-model for Africa country
mu.Africa <- apply (rugged.ensemble2$link , 2, mean )#based on averaged-model calculating mu 

# summarize to intervals
mu.Africa.PI <- apply( rugged.ensemble2$link , 2 , PI , prob=0.97 )

#plot the posterior predictions for The averaged-model for Africa country
d.A1 <- dd[dd$cont_africa==1,]
plot(log_gdp ~ rugged, data=d.A1,col=rangi2,ylab="log GDP 2000",xlab="Terrain Ruggedness Index")
mtext("Model-averaged predictions for Africa countrys ")
lines(rugged.seq,mu.Africa,col=rangi2)
shade(mu.Africa.PI,rugged.seq,col = col.alpha(rangi2,0.3))

#
```
###7H4
```{r}
library(rethinking)
##loads the data sets
data(nettle)
d <- nettle

#model the number of languages per capita as the outcome variable
d$lang.per.cap <- d$num.lang / d$k.pop

#(a)
# make log version of outcome and covariate
d$log_lang <- log( d$lang.per.cap )
d$log_area <- log(d$area)
summary(d)

##set the model for 7H4(a)
m.1.1 <- map(
  alist(
    log_lang ~ dnorm( mu , sigma ) ,
    mu <- a + bS*mean.growing.season ,
    a ~ dnorm( -6 , 100 ) ,
    bS ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=d )
precis(m.1.1)

plot(precis(m.1.1))

#plot postrior predictions
mean.growing.season.seq <- seq(from=-1,to=13,length.out = 50)
d.predict <- list(
  log_lang = rep(0,50),#empty outcome
  mean.growing.season = mean.growing.season.seq
)
pred.m.1.1 <- link(m.1.1,data=d.predict)
mu <- apply(pred.m.1.1,2, mean)
mu.PI <- apply(pred.m.1.1, 2, PI)

plot(log_lang ~ mean.growing.season,d,col=rangi2)
lines(mean.growing.season.seq,mu,col=rangi2)
shade(mu.PI,mean.growing.season.seq,col = col.alpha(rangi2,0.3))

#log_lang ~ Normal(mu, sigma)
#mu = a + bS*mean.growing.season + bA*log_area
m.1.2 <- map(
  alist(
    log_lang ~ dnorm( mu , sigma ) ,
    mu <- a + bS*mean.growing.season + bA*log_area,
    a ~ dnorm( -6 , 100 ) ,
    bS ~ dnorm( 0 , 1 ) ,
    bA ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=d )
precis(m.1.2)

#plot postrior predictions
mean.growing.season.seq <- seq(from=-1,to=13,length.out = 50)
d.predict <- list(
  log_lang = rep(0,50),#empty outcome
  mean.growing.season = mean.growing.season.seq,
  log_area = rep (13,50)#average log_area
)
pred.m.1.2 <- link(m.1.2,data=d.predict)
mu <- apply(pred.m.1.2,2, mean)
mu.PI <- apply(pred.m.1.2, 2, PI)

plot(log_lang ~ mean.growing.season,d,col=rangi2)
mtext("consider log(area) as a covariance")
lines(mean.growing.season.seq,mu,col=rangi2)
shade(mu.PI,mean.growing.season.seq,col = col.alpha(rangi2,0.3))

#compare two models
coeftab(m.1.1,m.1.2)
compare(m.1.1,m.1.2)

##set the model for 7H4(b)
##langrage diversity ~ standard deviation of length of growing season
m.2.1 <- map(
  alist(
    log_lang ~ dnorm( mu , sigma ) ,
    mu <- a + bS*sd.growing.season ,
    a ~ dnorm( -6 , 100 ) ,
    bS ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=d )
precis(m.2.1)

#plot postrior predictions
sd.growing.season.seq <- seq(from=-1,to=7,length.out = 50)
d.predict <- list(
  log_lang = rep(0,50),#empty outcome
  sd.growing.season = sd.growing.season.seq
)
pred.m.2.1 <- link(m.2.1,data=d.predict)
mu <- apply(pred.m.2.1,2, mean)
mu.PI <- apply(pred.m.2.1, 2, PI)

plot(log_lang ~ sd.growing.season,d,col=rangi2)
lines(sd.growing.season.seq,mu,col=rangi2)
shade(mu.PI,sd.growing.season.seq,col = col.alpha(rangi2,0.3))

##consider log area as a covariance
#log_lang ~ Normal(mu, sigma)
#mu = a + bS*sd.growing.season + bA*log_area
m.2.2 <- map(
  alist(
    log_lang ~ dnorm( mu , sigma ) ,
    mu <- a + bS*sd.growing.season + bA*log_area,
    a ~ dnorm( -6 , 100 ) ,
    bS ~ dnorm( 0 , 1 ) ,
    bA ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=d )
precis(m.2.2)

#plot postrior predictions
sd.growing.season.seq <- seq(from=-1,to=7,length.out = 50)
d.predict <- list(
  log_lang = rep(0,50),#empty outcome
  sd.growing.season = sd.growing.season.seq,
  log_area = rep (13,50)#average log_area
)
pred.m.2.2 <- link(m.2.2,data=d.predict)
mu <- apply(pred.m.2.2,2, mean)
mu.PI <- apply(pred.m.2.2, 2, PI)

plot(log_lang ~ sd.growing.season,d,col=rangi2)
mtext("consider log(area) as a covariance")
lines(sd.growing.season.seq,mu,col=rangi2)
shade(mu.PI,sd.growing.season.seq,col = col.alpha(rangi2,0.3))

#compare two models 
coeftab(m.2.1,m.2.2)
compare(m.2.1,m.2.2)

##set the model for 7H4(c)
m.3.1 <- map(
  alist(
    log_lang ~ dnorm( mu , sigma ) ,
    mu <- a + bM*mean.growing.season + bS*sd.growing.season + bMS*mean.growing.season*sd.growing.season ,
    a ~ dnorm( -6 , 100 ) ,
    bM ~ dnorm( 0 , 1 ) ,
    bS ~ dnorm( 0 , 1 ) ,
    bMS ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=d )
precis(m.3.1)

#consider log area as a covariance
#log_lang ~ Normal(mu, sigma)
#mu = a + bM*mean.growing.season + bS*sd.growing.season + bA*log_area
m.3.2 <- map(
  alist(
    log_lang ~ dnorm( mu , sigma ) ,
    mu <- a + bM*mean.growing.season + bS*sd.growing.season + bMS*mean.growing.season*sd.growing.season + bA*log_area,
    a ~ dnorm( -6 , 100 ) ,
    bM ~ dnorm( 0 , 1 ) ,
    bS ~ dnorm( 0 , 1 ) ,
    bA ~ dnorm( 0 , 1 ) ,
    bMS ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 10 )
  ) ,
  data=d )
precis(m.3.2)
#compare two models 
coeftab(m.3.1,m.3.2)
compare(m.3.1,m.3.2)

#averaged-model
models.ensemble <- ensemble(m.3.1,m.3.2,data=d)

#plot posterior predictions
mean.growing.season.seq <- seq(from=-1,to=13,length.out = 50)
sd.growing.season.seq <- seq(from=-1,to=7,length.out = 50)
#......
```

