# Statistical Rethinking Chapter 3 problems

__Name:__Xiaoyan Yue__


install.packages(c("coda","mvtnorm","devtools"))
library(devtools)
devtools::install_github("rmcelreath/rethinking")
library(rethinking)

```{r}
library(ggplot2)
library(devtools)
p_grid <- seq(from=0, to=1, length.out = 1000)
prior <- rep(1, 1000)
likelihood <- dbinom(6, size=9, prob=p_grid)
posterior <-likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample(p_grid, prob=posterior, size=1e4, replace = TRUE)
plot(samples)
qplot(samples)
dens(samples)
```
## 3E1 How much posterior probability lies below p = 0.2?
```{r}
sum(samples < 0.2)/1e4
```
## 3E2 How much posterior probability lies above p = 0.8?
```{r}
sum(samples > 0.8)/1e4
```
## 3E3 How much posterior probability lies between p = 0.2 and p = 0.8?
```{r}
sum(samples > 0.2 & samples < 0.8)/1e4
```
## 3E4 20% of the posterior probability lies below which value of p?
```{r}
quantile(samples, 0.2)
```
## 3E5 20% of the posterior probability lies above which value of p?  
```{r}
quantile(samples, 0.8) 
```
## 3E6 Which values of p contain the narrowest interval equal to 66% of the posterior probability? 
```{r}
HPDI(samples, prob = 0.66)
```

## 3E7 Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?
```{r}
PI(samples, prob=0.66)
```

## 3M1
```{r}
p_grid <- seq(from=0, to=1, length.out = 1000)
prior <- rep(1, 1000)
likelihood <- dbinom(8, size=15, prob=p_grid)
posterior <-likelihood * prior
posterior <- posterior / sum(posterior)
```
## 3M2
```{r}
p_grid <- seq(from=0, to=1, length.out = 1000)
prior <- rep(1, 1000)
likelihood <- dbinom(8, size=15, prob=p_grid)
posterior <-likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample(p_grid, prob=posterior, size=1e4, replace = TRUE)
HPDI(samples, 0.9)
```
_STOP AFTER 3M2 FOR 02/25 ASSIGNMENT_

## 3M3
```{r}
w <- rbinom(1e5, size=15,prob=samples)
sum(w==8)/1e4
```

## 3M4
```{r}
w <- rbinom(1e5, size=9,prob=samples)
sum(w==6)/1e5
```
## 3M5
####prior is zero below p=0.5 and a constant above p=0.5
```{r}
p_grid <- seq(from=0, to=1, length.out = 1000)
prior <- ifelse(p_grid < 0.5, 0, 1)
likelihood <- dbinom(8, size=15, prob=p_grid)
posterior <-likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
##simulate the distribution of samples
samples <- sample(p_grid, prob=posterior, size=1e4, replace = TRUE)
##what is the probility of observing 8 water in 15 tosses
w <- rbinom(1e5, size=15,prob=samples)
sum(w==8)/1e5
##what is probility of observing 6 water in 9 tosses
w <- rbinom(1e5, size=9,prob=samples)
sum(w==6)/1e5
```

## 3H1
library(rethinking)
data(homeworkch3)
sum(birth1) + sum(birth2)
```{R}
p_grid <- seq(from=0, to=1, length.out = 1000)
prior <-rep(1,1000)
likelihood <- dbinom(111, size=200, prob=p_grid)
posterior <-likelihood * prior
posterior <- posterior / sum(posterior)
posterior
p_grid[which.max(posterior)]
```

## 3H2
```{r}
samples <- sample(p_grid, prob=posterior, size=1e4, replace = TRUE)
HPDI(samples, prob = 0.5)
HPDI(sampels, prob = 0.89)
HPDI(samples, prob = 0.97)
```
## 3H3
```{r}
birth_boy <- rbinom(1e4,size=200,prob=samples)
simplehist( birth_boy , xlab="birth boy count" )
sum(birth_boy==111)/1e4
dens(birth_boy)
```
## 3H4
```{r}
sum(birth1)
p_grid <- seq(from=0, to=1, length.out = 1000)
prior <-rep(1,1000)
likelihood <- dbinom(51, size=100, prob=p_grid)
posterior <-likelihood * prior
posterior <- posterior / sum(posterior)
posterior
samples <- sample(p_grid, prob=posterior, size=1e4, replace = TRUE)
birth1_boy <- rbinom(1e4,size=100,prob=samples)
sum(birth1_boy==51)/1e4
dens(samples)
```
## 3H5